---
title: 'Table `1`: Descriptives of Index and Partner Cases'
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

# Objective 

The purpose here is to create a prototype of Table 1 - summarizing demographic
and behavioral characteristics of index persons and their contacts.



```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE,
                      root.dir = '~/code/rtp-network/out/')

library(data.table)
library(dplyr)
library(stringi)
library(network)
```



```{r read_data, include=FALSE}
data_dir <- "/gpfs/data/rkantor/rtp/datasets/D30_20211013_V1"
list.files(path=data_dir)
net_dt <- as.data.table(read.csv(paste0(data_dir, "/ContactTracingNetwork.csv")))
individuals_dt <- as.data.table(read.csv(paste0(data_dir, "/Individuals.csv")))


```

## Structure of the Partner Contact Database

```{r}
dim(net_dt)
```

The partner data are 4070 observations and 24 variables.

How many unique persons (sources of targets) are in the partner database?

```{r}
ind_in_partner_db <- unique(c(net_dt$StudyIDFrom, net_dt$StudyIDTo))
length(ind_in_partner_db)
```



How many unique index persons are in the partner database?

```{r}
# Identify all unique StudyIDFroms 
length(unique(net_dt$StudyIDFrom))

# Identify all unique StudyIDTos 
length(unique(net_dt$StudyIDTo))

# Identify the union of the two above
length(union(unique(net_dt$StudyIDFrom), unique(net_dt$StudyIDTo)))

# Identify their intersection
length(intersect(unique(net_dt$StudyIDFrom), unique(net_dt$StudyIDTo)))

# Test the cardinality of the above sets
len_studyidfrom <- length(unique(net_dt$StudyIDFrom))
len_studyidto <- length(unique(net_dt$StudyIDTo))
len_union <- length(union(unique(net_dt$StudyIDFrom), unique(net_dt$StudyIDTo)))
len_intersection <- length(intersect(unique(net_dt$StudyIDFrom), unique(net_dt$StudyIDTo)))

len_union == len_studyidfrom + len_studyidto - len_intersection

# Partition the data
unique_persons_dt <- unique(net_dt[,1, with=FALSE])
nrow(unique_persons_dt)

# Filter the data where StudyIDTo is not empty
net_dt_without_empty_StudyIDTo <- net_dt[StudyIDTo != ""]
nrow(net_dt_without_empty_StudyIDTo)

# Count the unique persons in StudyIDFrom column in the filtered dataset
length(unique(net_dt_without_empty_StudyIDTo$StudyIDFrom))

# Count the unique persons in StudyIDTo column in the filtered dataset
length(unique(net_dt_without_empty_StudyIDTo$StudyIDTo))
```

<span style="background-color: #FFFF00">
> The CTDB included disease intervention specialist interview data from `r length(length(ind_in_partner_db))` unique persons (`r length(unique(net_dt$StudyIDFrom))` newly diagnosed; `r length(unique(net_dt$StudyIDTo))` named partners). 
</span>

Number of unique persons in StudyIDFrom column: `r nrow(unique_persons_dt)`.

Number of unique StudyIDFrom's without an empty StudyIDTo: `r length(unique(net_dt_without_empty_StudyIDTo$StudyIDFrom))`. 

When were the data collected?

```{r}
table(substr(net_dt$IdentifiedDate, 1, 4), exclude = NULL)
```







## Structure of the Statewide Unified Individuals' Genomic Database 

How many unique persons in the individuals database?

```{r}

head(individuals_dt)

nrow(individuals_dt)
length(unique(individuals_dt$StudyID))

table(substr(individuals_dt$StudyID, 1, 1))
```
**Note**: There are no persons with "CONTACT" Study ID in the unified individuals
database. 

What is the source of the information of persons in the individuals database?

```{r}
table(individuals_dt$Source, exclude = NULL)
table(individuals_dt$Source, exclude = NULL)/sum(table(individuals_dt$Source, exclude = NULL))
```


How many unique clusters are the individuals in? 
```{r}
length(unique(individuals_dt$ClusteredHIVTrace005)) 
length(unique(individuals_dt$ClusteredHIVTrace015)) 
length(unique(individuals_dt$ClusteredPhyloAny)) 
```




## Intersection of the Two Databases


How many individuals from the genomics database appear in the partner database,
as StudyIDFrom, and StudyIDTo?

```{r}
individuals_in_net_dt_studyidfrom <- 
  which(individuals_dt$StudyID %in% net_dt$StudyIDFrom)
length(individuals_in_net_dt_studyidfrom)

individuals_in_net_dt_studyidto <- which(individuals_dt$StudyID %in% net_dt$StudyIDTo)
length(individuals_in_net_dt_studyidto)

individuals_in_net_dt_studyidfrom_or_to <- 
  which(individuals_dt$StudyID %in% c(net_dt$StudyIDFrom, net_dt$StudyIDTo))
length(individuals_in_net_dt_studyidfrom_or_to)

individuals_in_net_dt_studyidto <- which(individuals_dt$StudyID %in% net_dt$StudyIDTo)
length(individuals_in_net_dt_studyidto)

individuals_in_net_dt_studyidfrom_wo_emptyStudyIDTo <- 
  which(individuals_dt$StudyID %in% net_dt_without_empty_StudyIDTo$StudyIDFrom)
length(individuals_in_net_dt_studyidfrom_wo_emptyStudyIDTo)

individuals_in_net_dt_studyidto_witout_empty_StudyIDTo <- 
  which(individuals_dt$StudyID %in% net_dt_without_empty_StudyIDTo$StudyIDTo)
length(individuals_in_net_dt_studyidto_witout_empty_StudyIDTo)
```


- There are `r length(unique(individuals_dt[individuals_in_net_dt_studyidfrom,]$StudyID))` unique individuals in the Unified Individuals dataset who appear in the Partner 
Contact database (including with empty StudyIDTos).

- There are  `r length(unique(individuals_in_net_dt_studyidfrom_wo_emptyStudyIDTo))`
unique individuals in the Unified Individuals dataset who appear in the Partner 
Contact database (without any empty StudyIDTos, i.e. they reported on at least 
one partner).

- There are `r length(unique(individuals_in_net_dt_studyidto))` reported 
unique partners of persons in the Unified Individuals dataset who appear in the 
partner contact dataset (including StudyIDTos with empty strings).

- There are `r length(unique(individuals_in_net_dt_studyidto_witout_empty_StudyIDTo))` reported 
unique partners of persons in the Unified Individuals dataset who appear in the 
Partner Contact dataset  (excluding StudyIDTos with empty strings).

#### How many persons for whom we have sequences appear as both 
index persons and partners?

```{r}
intersection_set_from_and_to <- sort(
  intersect(
  individuals_in_net_dt_studyidfrom_wo_emptyStudyIDTo,
  individuals_in_net_dt_studyidto_witout_empty_StudyIDTo
))
  
length(intersection_set_from_and_to)
```

Of the newly diagnosed persons in the contact tracing database, 
for how many do we have sequences and how many are in clusters? 
How many are clustered with someone who named them?

```{r}
table(individuals_dt[individuals_in_net_dt_studyidfrom_wo_emptyStudyIDTo]$ClusteredPhyloAny, 
      exclude = NULL)
sum(table(individuals_dt[individuals_in_net_dt_studyidfrom_wo_emptyStudyIDTo]$ClusteredPhyloAny, 
      exclude = NULL))
```

Ans: 

* Number of newly diagnosed persons in the contact tracing database = 
  `r length(individuals_in_net_dt_studyidfrom)`
  
* Of these, `r sum(table(individuals_dt[individuals_in_net_dt_studyidfrom_wo_emptyStudyIDTo]$ClusteredPhyloAny, exclude = NULL))` persons provided partner data. 

* Of these, `r length(which(!is.na(individuals_dt[individuals_in_net_dt_studyidfrom_wo_emptyStudyIDTo]$ClusteredPhyloAny)))` are clustered phylogenetically.

Of the partners of the newly diagnosed persons in the contact tracing database, 
how many nominated other persons? How many are clustered with someone who named them?

```{r}
table(individuals_dt[individuals_in_net_dt_studyidto_witout_empty_StudyIDTo]$ClusteredPhyloAny, 
      exclude = NULL)
sum(table(individuals_dt[individuals_in_net_dt_studyidto_witout_empty_StudyIDTo]$ClusteredPhyloAny, 
      exclude = NULL))
```


Ans:   

* Of the partners of the newly diagnosed persons in the contact tracing database, 
`r sum(table(individuals_dt[individuals_in_net_dt_studyidto_witout_empty_StudyIDTo]$ClusteredPhyloAny, exclude = NULL))` nominated other persons.

* Of these, `r length(which(!is.na(individuals_dt[individuals_in_net_dt_studyidto_witout_empty_StudyIDTo]$ClusteredPhyloAny)))` clustered phylogenetically with someone who named them.


What is the Jaccard Similarity Coefficient for the individuals database and partner contact tracing database?

```{r}
study_ids_ctdb <- unique(c(net_dt$StudyIDFrom, net_dt$StudyIDTo))
study_ids_gdb <- individuals_dt$StudyID

intersection <- length(intersect(study_ids_ctdb, study_ids_gdb))
union <- length(unique(c(study_ids_ctdb, study_ids_gdb)))

jaccard_similarity <- intersection / union

```

The Jaccard Similarity Index is `r jaccard_similarity`.

For how many persons (reported as
index cases or partners in the PCRS dataset) do we have  
sequences? 

```{r}
study_id_from_and_to_in_individuals_dt <- 
  sort(union(
  individuals_in_net_dt_studyidfrom_wo_emptyStudyIDTo,
  individuals_in_net_dt_studyidto_witout_empty_StudyIDTo
  )
  )
which(duplicated(study_id_from_and_to_in_individuals_dt) == TRUE)
length(study_id_from_and_to_in_individuals_dt)
```

- There are `r length(unique(study_id_from_and_to_in_individuals_dt))` unique
persons (reported as
index cases or partners in the PCRS dataset) for whom we have  
sequences.

Note that these `r length(unique(study_id_from_and_to_in_individuals_dt))`  are comprised of 
`r length(individuals_in_net_dt_studyidfrom_wo_emptyStudyIDTo)` 
who appear as index cases who are interviewed for PCRS, 
`r length(individuals_in_net_dt_studyidto_witout_empty_StudyIDTo)` 
appeared as reported partners and are in the genomics database, and 
`r length(intersection_set_from_and_to)`
 appeared as both index persons and partners. 

Note that 
```{r}
identical(length(unique(study_id_from_and_to_in_individuals_dt)),
           (length(individuals_in_net_dt_studyidfrom_wo_emptyStudyIDTo) +
              length(individuals_in_net_dt_studyidto_witout_empty_StudyIDTo) -
              length(intersection_set_from_and_to)
            )
)
```

567 = 509 + 156 – 98. 

#### How many persons are in the genomic database but not in the PCRS database?

```{r}

length(which(individuals_dt$StudyID %in% ind_in_partner_db == FALSE))
```


#### How many persons are in the PCRS database but not in the genomic database?

```{r}
length(which(ind_in_partner_db %in% individuals_dt$StudyID == FALSE))
```


#### What are the cluster sizes per the three sequencing techniques 
of persons who appear in the partner database? 



```{r}
length(unique(individuals_dt$ClusteredHIVTrace005[study_id_from_and_to_in_individuals_dt])) 
length(unique(individuals_dt$ClusteredHIVTrace015[study_id_from_and_to_in_individuals_dt])) 
length(unique(individuals_dt$ClusteredPhyloAny[study_id_from_and_to_in_individuals_dt])) 
```

## [Table "1"](https://docs.google.com/document/d/12g9p5gsKfzjGl-hP524krx7GwZPHGY1UBvMf_7cMo8I/edit#bookmark=id.tc8pri6l7bee)  
**Descriptive Characteristics of Index Persons and Partner Contacts**


#### By source

```{r, include=TRUE, echo=TRUE}

  individuals_dt[individuals_in_net_dt_studyidfrom, .(n_source=.N), 
               by=.(Source)
               ][,
                 "prop":=round(n_source/sum(n_source), 3)
                 ][
                 ]


  individuals_dt[individuals_in_net_dt_studyidfrom_wo_emptyStudyIDTo, .(n_source=.N), 
               by=.(Source)
               ][,
                 "prop":=round(n_source/sum(n_source), 3)
                 ][
                 ]

individuals_dt[individuals_in_net_dt_studyidto, .(n_source=.N), 
               by=.(Source)
               ][,
                 "prop":=round(n_source/sum(n_source), 3)
                 ][
                 ]

```

#### Gender and Birth Sex

```{r, include=TRUE, echo=TRUE}
individuals_dt[individuals_in_net_dt_studyidfrom, .(.N), 
               by=.(DemoGender)
               ][,
                 "prop":=round(N/sum(N), 3)
                 ][
                 ]

individuals_dt[individuals_in_net_dt_studyidfrom, .(.N), 
               by=.(DemoBirthSex)
               ][,
                 "prop":=round(N/sum(N), 3)
                 ][
                 ]

individuals_dt[individuals_in_net_dt_studyidfrom_wo_emptyStudyIDTo, .(.N), 
               by=.(DemoGender)
               ][,
                 "prop":=round(N/sum(N), 3)
                 ][
                 ]

individuals_dt[individuals_in_net_dt_studyidfrom_wo_emptyStudyIDTo, .(.N), 
               by=.(DemoBirthSex)
               ][,
                 "prop":=round(N/sum(N), 3)
                 ][
                 ]


individuals_dt[individuals_in_net_dt_studyidto, .(.N), 
               by=.(DemoGender)
               ][,
                 "prop":=round(N/sum(N), 3)
                 ][
                 ]

individuals_dt[individuals_in_net_dt_studyidto, .(.N), 
               by=.(DemoBirthSex)
               ][,
                 "prop":=round(N/sum(N), 3)
                 ][
                 ]
```

#### Race and Ethnicity

```{r, include=TRUE, echo=TRUE}
individuals_dt[individuals_in_net_dt_studyidfrom, .(.N), 
               by=.(DemoRace)
               ][,
                 "prop":=round(N/sum(N), 3)
                 ][
                 ]

individuals_dt[individuals_in_net_dt_studyidfrom, .(.N), 
               by=.(DemoHispanic)
               ][,
                 "prop":=round(N/sum(N), 3)
                 ][
                 ]

individuals_dt[individuals_in_net_dt_studyidfrom_wo_emptyStudyIDTo, .(.N), 
               by=.(DemoRace)
               ][,
                 "prop":=round(N/sum(N), 3)
                 ][]

individuals_dt[individuals_in_net_dt_studyidfrom_wo_emptyStudyIDTo, .(.N), 
               by=.(DemoHispanic)
               ][,
                 "prop":=round(N/sum(N), 3)
                 ][]

individuals_dt[individuals_in_net_dt_studyidto, .(.N), 
               by=.(DemoRace)
               ][,
                 "prop":=round(N/sum(N), 3)
                 ][
                 ]

individuals_dt[individuals_in_net_dt_studyidto, .(.N), 
               by=.(DemoHispanic)
               ][,
                 "prop":=round(N/sum(N), 3)
                 ][
                 ]

```


#### Age at HIV Diagnosis

**Index Persons** 

```{r}
individuals_dt[individuals_in_net_dt_studyidfrom, 
               .(mean=mean(HIVDxAge, na.rm=T), 
                 median=median(HIVDxAge, na.rm = T), 
                 sd=sd(HIVDxAge, na.rm = T), 
                 min=min(HIVDxAge, na.rm = TRUE), 
                 max=max(HIVDxAge, na.rm=TRUE))
               ][
                 ]

## discretized
breaks = c(0, 19, 25, 35, 45, 55, 100)
labels = c("<=18", "19-24", "25-34", "35-44", "45-54", ">=55")

individuals_dt[individuals_in_net_dt_studyidfrom, 
               hivdxage_cat := 
                        cut(HIVDxAge, breaks = breaks, 
                            labels = labels)]

hivdxage_cat_dt <- 
  individuals_dt[individuals_in_net_dt_studyidfrom, 
               .(.N), by=hivdxage_cat][,
                 "prop":=round(N/sum(N), 3)
                 ][order(hivdxage_cat)]
hivdxage_cat_dt
sum(hivdxage_cat_dt$N)

## Who are the people here who do not have a reported age for diagnosis?
na_hivdxage_cat_dt <- individuals_dt[intersect(individuals_in_net_dt_studyidfrom, which(is.na(individuals_dt$hivdxage_cat))),]
dim(na_hivdxage_cat_dt)
knitr::kable(na_hivdxage_cat_dt)
knitr::kable(individuals_dt[individuals_in_net_dt_studyidfrom, .SD, .SDcols=c("StudyID", "Source", "HIVDxDate", "HIVDxAge", "hivdxage_cat")])

```

**Index Persons (reporting at least one partner)** 

```{r}
individuals_dt[individuals_in_net_dt_studyidfrom_wo_emptyStudyIDTo, 
               .(mean=mean(HIVDxAge, na.rm=T), 
                 median=median(HIVDxAge, na.rm = T), 
                 sd=sd(HIVDxAge, na.rm = T), 
                 min=min(HIVDxAge, na.rm = TRUE), 
                 max=max(HIVDxAge, na.rm=TRUE))
               ][
                 ]

## discretized
breaks = c(0, 19, 25, 35, 45, 55, 100)
labels = c("<=18", "19-24", "25-34", "35-44", "45-54", ">=55")

individuals_dt[individuals_in_net_dt_studyidfrom_wo_emptyStudyIDTo, 
               hivdxage_cat := 
                        cut(HIVDxAge, breaks = breaks, 
                            labels = labels)]

hivdxage_cat_dt <- 
  individuals_dt[individuals_in_net_dt_studyidfrom_wo_emptyStudyIDTo, 
               .(.N), by=hivdxage_cat][,
                 "prop":=round(N/sum(N), 3)
                 ][order(hivdxage_cat)]
hivdxage_cat_dt
sum(hivdxage_cat_dt$N)

## Who are the people here who do not have a reported age for diagnosis?
na_hivdxage_cat_dt <- individuals_dt[intersect(individuals_in_net_dt_studyidfrom_wo_emptyStudyIDTo, which(is.na(individuals_dt$hivdxage_cat))),]
dim(na_hivdxage_cat_dt)
knitr::kable(na_hivdxage_cat_dt)
knitr::kable(individuals_dt[individuals_in_net_dt_studyidfrom, .SD, .SDcols=c("StudyID", "Source", "HIVDxDate", "HIVDxAge", "hivdxage_cat")])

```

**Partner Contacts**

```{r}
individuals_dt[individuals_in_net_dt_studyidto, 
               .(mean=mean(HIVDxAge, na.rm=T), 
                 median=median(HIVDxAge, na.rm = T), 
                 sd=sd(HIVDxAge, na.rm = T), 
                 min=min(HIVDxAge, na.rm = TRUE), 
                 max=max(HIVDxAge, na.rm=TRUE))
               ][
                 ]

## discretized
breaks = c(0, 19, 25, 35, 45, 55, 100)
labels = c("<=18", "19-24", "25-34", "35-44", "45-54", ">=55")

individuals_dt[individuals_in_net_dt_studyidto, 
               hivdxage_cat := 
                        cut(HIVDxAge, breaks = breaks, 
                            labels = labels)]

hivdxage_cat_dt <- 
  individuals_dt[individuals_in_net_dt_studyidto, 
               .(.N), by=hivdxage_cat][,
                 "prop":=round(N/sum(N), 3)
                 ][order(hivdxage_cat)]
hivdxage_cat_dt
sum(hivdxage_cat_dt$N)

## Who are the people here who do not have a reported age for diagnosis?
na_hivdxage_cat_dt <- individuals_dt[intersect(individuals_in_net_dt_studyidto, which(is.na(individuals_dt$hivdxage_cat))),]
dim(na_hivdxage_cat_dt)
knitr::kable(na_hivdxage_cat_dt)
knitr::kable(individuals_dt[individuals_in_net_dt_studyidto, .SD, .SDcols=c("StudyID", "Source", "HIVDxDate", "HIVDxAge", "hivdxage_cat")])

```

```{r}
individuals_dt[individuals_in_net_dt_studyidfrom, lapply(.SD, function (x) length(x)), by=Source]
```

#### HIV Diagnosis Year

**Index Persons**
```{r}
index_persons_dt <- individuals_dt[individuals_in_net_dt_studyidfrom,]
index_persons_dt[, hivdiagnosis_year := as.numeric(substr(HIVDxDate, 1, 4))]
table(index_persons_dt$hivdiagnosis_year)

breaks = c(0, 2000, 2006, 2011, 2021)
labels = c("<2000", "2001-2005", "2006-2010", "2011-2020")

index_persons_dt[, hivdiagnosis_year_group := 
                        cut(hivdiagnosis_year, breaks = breaks, 
                            labels = labels)]
table(index_persons_dt$hivdiagnosis_year_group, exclude = NULL)

index_persons_dt[, .(.N), by=hivdiagnosis_year_group][,
                 "prop":=round(N/sum(N), 3)
                 ][order(hivdiagnosis_year_group)]
```

```{r}
index_persons_dt[which(is.na(hivdiagnosis_year)), .SD, .SDcols=c("StudyID", "Source", "DemoGender", "HIVDxDate", "HIVDxAge", "hivdiagnosis_year", "hivdiagnosis_year_group")]
```

There are 40 index persons who do not have a reported year of HIV Diagnosis. 
All of these are persons whose source is RIDOH. 

** Index Persons (reporting at least one contact) **

```{r}
index_persons_ge1contact_dt <- 
  individuals_dt[individuals_in_net_dt_studyidfrom_wo_emptyStudyIDTo,]
index_persons_ge1contact_dt[, hivdiagnosis_year := 
                              as.numeric(substr(HIVDxDate, 1, 4))]
table(index_persons_ge1contact_dt$hivdiagnosis_year)

breaks = c(0, 2000, 2006, 2011, 2021)
labels = c("<2000", "2001-2005", "2006-2010", "2011-2020")

index_persons_ge1contact_dt[, hivdiagnosis_year_group := 
                        cut(hivdiagnosis_year, breaks = breaks, 
                            labels = labels)]
table(index_persons_ge1contact_dt$hivdiagnosis_year_group, exclude = NULL)

index_persons_ge1contact_dt[, .(.N), 
                            by=hivdiagnosis_year_group][,
                 "prop":=round(N/sum(N), 3)
                 ][order(hivdiagnosis_year_group)]
```

**Partner Contacts**
```{r}
partner_contacts_dt <- individuals_dt[individuals_in_net_dt_studyidto,]
partner_contacts_dt[, hivdiagnosis_year := as.numeric(substr(HIVDxDate, 1, 4))]
table(partner_contacts_dt$hivdiagnosis_year)

breaks = c(0, 2000, 2006, 2011, 2021)
labels = c("<2000", "2001-2005", "2006-2010", "2011-2020")

partner_contacts_dt[, hivdiagnosis_year_group := 
                        cut(hivdiagnosis_year, breaks = breaks, 
                            labels = labels)]
table(partner_contacts_dt$hivdiagnosis_year_group, exclude = NULL)

partner_contacts_dt[, .(.N), by=hivdiagnosis_year_group][,
                 "prop":=round(N/sum(N), 3)
                 ][order(hivdiagnosis_year_group)]
```


#### Behavioral Characteristics: MSM, IDU, HRH

**Index Persons**
```{r}
index_persons_dt[, .(.N), by=RiskMSM][,
                 "prop":=round(N/sum(N), 3)
                 ][]

index_persons_dt[, .(.N), by=RiskIDU][,
                 "prop":=round(N/sum(N), 3)
                 ][]

index_persons_dt[, .(.N), by=RiskHRH][,
                 "prop":=round(N/sum(N), 3)
                 ][]
```


**Index Persons (reporting at least one partner)**
```{r}
index_persons_ge1contact_dt[, .(.N), by=RiskMSM][,
                 "prop":=round(N/sum(N), 3)
                 ][]

index_persons_ge1contact_dt[, .(.N), by=RiskIDU][,
                 "prop":=round(N/sum(N), 3)
                 ][]

index_persons_ge1contact_dt[, .(.N), by=RiskHRH][,
                 "prop":=round(N/sum(N), 3)
                 ][]
```


**Partner Contacts**
```{r}
partner_contacts_dt[, .(.N), by=RiskMSM][,
                 "prop":=round(N/sum(N), 3)
                 ][]

partner_contacts_dt[, .(.N), by=RiskIDU][,
                 "prop":=round(N/sum(N), 3)
                 ][]

partner_contacts_dt[, .(.N), by=RiskHRH][,
                 "prop":=round(N/sum(N), 3)
                 ][]
```

### [Table 2]()  
**Mean number of named, diagnosed and sequenced partners by index case characteristics**

```{r}
#knitr::kable(net_dt)
```

To focus on index persons with a named person, 
filter out rows with missing `StudyIDTo`:

```{r}
no_missing_rows_in_net_dt <- net_dt[StudyIDTo != "", , ]
class(no_missing_rows_in_net_dt)
```



- Start by computing the mean number of *named partners* for all index persons?
- Then break down this list by index person behavioral and demographic characteristics?

Number of contacts reported by each index person: 


#### Analysis of named partners for all index persons

```{r}
net_dt[, .N, by = StudyIDFrom] #full dataset
#net_dt[StudyIDFrom == "S20000441296", ]

no_missing_rows_in_net_dt[, .N, by = StudyIDFrom] #IPs with named partners only
#no_missing_rows_in_net_dt[StudyIDFrom == "S20000441296", ]
```

Number of unique contacts reported by each index person:

```{r}
net_dt[, .(UniqueContacts = uniqueN(StudyIDTo)), by = StudyIDFrom]

unique_contacts_counts <- no_missing_rows_in_net_dt[, .(UniqueContacts = uniqueN(StudyIDTo)), by = StudyIDFrom]

sum(unique_contacts_counts$UniqueContacts)
```
Count the number of times each contact has been reported:

```{r}
summary(as.numeric(table(net_dt$StudyIDTo)[-1]))

summary(as.numeric(table(no_missing_rows_in_net_dt$StudyIDTo)))
```

```{r}
individuals_in_net_dt_studyidfrom_nomiss <- which(individuals_dt$StudyID %in% no_missing_rows_in_net_dt$StudyIDFrom)
length(individuals_in_net_dt_studyidfrom_nomiss)

individuals_in_net_dt_studyidto_nomiss <- which(individuals_dt$StudyID %in% no_missing_rows_in_net_dt$StudyIDTo)
length(individuals_in_net_dt_studyidto_nomiss)
```

How many unique partners are reported by indexed persons with named partners?

```{r}
length(no_missing_rows_in_net_dt$StudyIDTo)
length(unique(no_missing_rows_in_net_dt$StudyIDTo))
```

What is the distribution of unique partners reported by indexed persons with named partners?

```{r}
pts_reported_by_IPs_with_named_pts <- no_missing_rows_in_net_dt[,.N,by=StudyIDFrom]
dim(pts_reported_by_IPs_with_named_pts)
summary(pts_reported_by_IPs_with_named_pts$N)

idxs <- which(pts_reported_by_IPs_with_named_pts$StudyIDFrom %in% individuals_dt$StudyID == TRUE)
IPs_reporting_partners_in_unified_dt <- pts_reported_by_IPs_with_named_pts$StudyIDFrom[idxs]
```

Of these `length(unique(no_missing_rows_in_net_dt$StudyIDTo))`, how many appear
in the unified individuals data set and how many are diagnosed?

**N.B.**: Date of HIV Diagnosis is used to indicate whether a person is diagnosed.

```{r}
unique_named_pt_contacts_reported_by_IPs <- unique(no_missing_rows_in_net_dt$StudyIDTo)

which(unique_named_pt_contacts_reported_by_IPs %in% individuals_dt$StudyID)

named_contacts_in_unified_dataset <- unique_named_pt_contacts_reported_by_IPs[which(unique_named_pt_contacts_reported_by_IPs %in% individuals_dt$StudyID)]


named_contacts_rows_from_unified_dt <- individuals_dt[StudyID %in% named_contacts_in_unified_dataset,,]
named_contacts_rows_from_unified_dt$HIVDxDate
length(named_contacts_rows_from_unified_dt$HIVDxDate)
length(which(named_contacts_rows_from_unified_dt$HIVDxDate == ""))
table(individuals_dt[StudyID %in% named_contacts_in_unified_dataset,,]$HIVDxDate, exclude = NULL)
```

How many are sequenced?

```{r}
table(named_contacts_rows_from_unified_dt$Sequence, exclude = NULL)
```

#### Analysis of named partners for index persons stratified by behavior,
race, ethnicity.

**MSM**:

Number of index persons (with named partners) who identify as MSM:

```{r}
individuals_dt_of_msm_who_named_partners <-
  individuals_dt[StudyID %in% IPs_reporting_partners_in_unified_dt & RiskMSM == "True",,]
nrow(individuals_dt_of_msm_who_named_partners)
```

***How many partners are reported by the `r nrow(individuals_dt_of_msm_who_named_partners)` MSM persons who named partners?***

```{r}
net_dt_of_msm_who_named_partners <- net_dt[StudyIDFrom %in% individuals_dt_of_msm_who_named_partners$StudyID,,] 

pt_counts_msm_who_named_partners <- net_dt_of_msm_who_named_partners[,.(N=.N),by=StudyIDFrom]

nrow(pt_counts_msm_who_named_partners)
pt_counts_msm_who_named_partners[,sum(N)]
pt_counts_msm_who_named_partners[,summary(N)]
```

***Ans: `r pt_counts_msm_who_named_partners[,sum(N)]`***.

***How many of these partners are in the individuals dataset, diagnosed, and sequenced?***

```{r}
idx <- which(net_dt_of_msm_who_named_partners$StudyIDTo %in% individuals_dt$StudyID)
msm_reported_partners_in_individuals_dt <- net_dt_of_msm_who_named_partners$StudyIDTo[idx] 
unique(msm_reported_partners_in_individuals_dt)

msm_partners_in_individuals_full_dt <- individuals_dt[StudyID %in% msm_reported_partners_in_individuals_dt,,]
table(msm_partners_in_individuals_full_dt$HIVDxDate, exclude = NULL)
table(msm_partners_in_individuals_full_dt$Sequence, exclude = NULL)
```


**IDU**:

Number of index persons (with named partners) who identify as IDU:

```{r}
individuals_dt_of_hrh_who_named_partners <-
  individuals_dt[StudyID %in% IPs_reporting_partners_in_unified_dt & RiskIDU == "True",,]
nrow(individuals_dt_of_hrh_who_named_partners)
```


***How many partners are reported by the `r nrow(individuals_dt_of_hrh_who_named_partners)` idu persons who named partners?***

```{r}
net_dt_of_hrh_who_named_partners <- net_dt[StudyIDFrom %in% individuals_dt_of_hrh_who_named_partners$StudyID,,] 

pt_counts_hrh_who_named_partners <- net_dt_of_hrh_who_named_partners[,.(N=.N),by=StudyIDFrom]

nrow(pt_counts_hrh_who_named_partners)
pt_counts_hrh_who_named_partners[,sum(N)]
pt_counts_hrh_who_named_partners[,summary(N)]
```

***Ans: `r pt_counts_hrh_who_named_partners[,sum(N)]`***.

***How many of these partners are in the individuals dataset, diagnosed, and sequenced?***

```{r}
idx <- which(net_dt_of_hrh_who_named_partners$StudyIDTo %in% individuals_dt$StudyID)
idx
idu_reported_partners_in_individuals_dt <- net_dt_of_hrh_who_named_partners$StudyIDTo[idx] 
idu_reported_partners_in_individuals_dt
unique(idu_reported_partners_in_individuals_dt)
length(unique(idu_reported_partners_in_individuals_dt))

idu_partners_in_individuals_full_dt <- individuals_dt[StudyID %in% idu_reported_partners_in_individuals_dt,,]
table(idu_partners_in_individuals_full_dt$HIVDxDate, exclude = NULL)
table(idu_partners_in_individuals_full_dt$Sequence, exclude = NULL)
```




**HRH**:

Number of index persons (with named partners) who identify as HRH:

```{r}
individuals_dt_of_hrh_who_named_partners <-
  individuals_dt[StudyID %in% IPs_reporting_partners_in_unified_dt & RiskHRH == "True",,]
nrow(individuals_dt_of_hrh_who_named_partners)
```


***How many partners are reported by the `r nrow(individuals_dt_of_hrh_who_named_partners)` idu persons who named partners?***

```{r}
net_dt_of_hrh_who_named_partners <- net_dt[StudyIDFrom %in% individuals_dt_of_hrh_who_named_partners$StudyID,,] 

pt_counts_hrh_who_named_partners <- net_dt_of_hrh_who_named_partners[,.(N=.N),by=StudyIDFrom]

nrow(pt_counts_hrh_who_named_partners)
pt_counts_hrh_who_named_partners[,sum(N)]
pt_counts_hrh_who_named_partners[,summary(N)]
```

***Ans: `r pt_counts_hrh_who_named_partners[,sum(N)]`***.

***How many of these partners are in the individuals dataset, diagnosed, and sequenced?***

```{r}
idx <- which(net_dt_of_hrh_who_named_partners$StudyIDTo %in% individuals_dt$StudyID)
idx
hrh_reported_partners_in_individuals_dt <- net_dt_of_hrh_who_named_partners$StudyIDTo[idx] 
hrh_reported_partners_in_individuals_dt
unique(hrh_reported_partners_in_individuals_dt)
length(unique(hrh_reported_partners_in_individuals_dt))

hrh_partners_in_individuals_full_dt <- individuals_dt[StudyID %in% hrh_reported_partners_in_individuals_dt,,]
table(hrh_partners_in_individuals_full_dt$HIVDxDate, exclude = NULL)
table(hrh_partners_in_individuals_full_dt$HIVDxDate, exclude = NULL)
table(hrh_partners_in_individuals_full_dt$Sequence, exclude = NULL)
```

**BLACK**:

Number of index persons (with named partners) who identify as Black:

```{r}
individuals_dt_of_black_who_named_partners <-
  individuals_dt[StudyID %in% IPs_reporting_partners_in_unified_dt & DemoRace == "Black",,]
nrow(individuals_dt_of_black_who_named_partners)
```


***How many partners are reported by the `r nrow(individuals_dt_of_black_who_named_partners)` Black persons who named partners?***

```{r}
net_dt_of_black_who_named_partners <- net_dt[StudyIDFrom %in% individuals_dt_of_black_who_named_partners$StudyID,,] 

pt_counts_black_who_named_partners <- net_dt_of_black_who_named_partners[,.(N=.N),by=StudyIDFrom]

nrow(pt_counts_black_who_named_partners)
pt_counts_black_who_named_partners[,sum(N)]
pt_counts_black_who_named_partners[,summary(N)]
```

***Ans: `r pt_counts_black_who_named_partners[,sum(N)]`***.

***How many of these partners are in the individuals dataset, diagnosed, and sequenced?***

```{r}
idx <- which(net_dt_of_black_who_named_partners$StudyIDTo %in% individuals_dt$StudyID)
idx
black_reported_partners_in_individuals_dt <- net_dt_of_black_who_named_partners$StudyIDTo[idx] 
black_reported_partners_in_individuals_dt
unique(black_reported_partners_in_individuals_dt)
length(unique(black_reported_partners_in_individuals_dt))

black_partners_in_individuals_full_dt <- individuals_dt[StudyID %in% black_reported_partners_in_individuals_dt,,]
table(black_partners_in_individuals_full_dt$HIVDxDate, exclude = NULL)
table(black_partners_in_individuals_full_dt$HIVDxDate, exclude = NULL)
table(black_partners_in_individuals_full_dt$Sequence, exclude = NULL)
```

**White**

Number of index persons (with named partners) who identify as White:

```{r}
individuals_dt_of_white_who_named_partners <-
  individuals_dt[StudyID %in% IPs_reporting_partners_in_unified_dt & DemoRace == "White",,]
nrow(individuals_dt_of_white_who_named_partners)
```


***How many partners are reported by the `r nrow(individuals_dt_of_white_who_named_partners)` White persons who named partners?***

```{r}
net_dt_of_white_who_named_partners <- net_dt[StudyIDFrom %in% individuals_dt_of_white_who_named_partners$StudyID,,] 

pt_counts_white_who_named_partners <- net_dt_of_white_who_named_partners[,.(N=.N),by=StudyIDFrom]

nrow(pt_counts_white_who_named_partners)
pt_counts_white_who_named_partners[,sum(N)]
pt_counts_white_who_named_partners[,summary(N)]
```

***Ans: `r pt_counts_white_who_named_partners[,sum(N)]`***.

***How many of these partners are in the individuals dataset, diagnosed, and sequenced?***

```{r}
idx <- which(net_dt_of_white_who_named_partners$StudyIDTo %in% individuals_dt$StudyID)
idx
white_reported_partners_in_individuals_dt <- net_dt_of_white_who_named_partners$StudyIDTo[idx] 
white_reported_partners_in_individuals_dt
unique(white_reported_partners_in_individuals_dt)
length(unique(white_reported_partners_in_individuals_dt))

white_partners_in_individuals_full_dt <- individuals_dt[StudyID %in% white_reported_partners_in_individuals_dt,,]
table(white_partners_in_individuals_full_dt$HIVDxDate, exclude = NULL)
table(white_partners_in_individuals_full_dt$HIVDxDate, exclude = NULL)
table(white_partners_in_individuals_full_dt$Sequence, exclude = NULL)
```

**Asian**

Number of index persons (with named partners) who identify as Asian:

```{r}
individuals_dt_of_asian_who_named_partners <-
  individuals_dt[StudyID %in% IPs_reporting_partners_in_unified_dt & DemoRace == "Asian",,]
nrow(individuals_dt_of_asian_who_named_partners)
```


***How many partners are reported by the `r nrow(individuals_dt_of_asian_who_named_partners)` Asian persons who named partners?***

```{r}
net_dt_of_asian_who_named_partners <- net_dt[StudyIDFrom %in% individuals_dt_of_asian_who_named_partners$StudyID,,] 

pt_counts_asian_who_named_partners <- net_dt_of_asian_who_named_partners[,.(N=.N),by=StudyIDFrom]

nrow(pt_counts_asian_who_named_partners)
pt_counts_asian_who_named_partners[,sum(N)]
pt_counts_asian_who_named_partners[,summary(N)]
```

***Ans: `r pt_counts_asian_who_named_partners[,sum(N)]`***.

***How many of these partners are in the individuals dataset, diagnosed, and sequenced?***

```{r}
idx <- which(net_dt_of_asian_who_named_partners$StudyIDTo %in% individuals_dt$StudyID)
idx
asian_reported_partners_in_individuals_dt <- net_dt_of_asian_who_named_partners$StudyIDTo[idx] 
asian_reported_partners_in_individuals_dt
unique(asian_reported_partners_in_individuals_dt)
length(unique(asian_reported_partners_in_individuals_dt))

asian_partners_in_individuals_full_dt <- individuals_dt[StudyID %in% asian_reported_partners_in_individuals_dt,,]
table(asian_partners_in_individuals_full_dt$HIVDxDate, exclude = NULL)
table(asian_partners_in_individuals_full_dt$HIVDxDate, exclude = NULL)
table(asian_partners_in_individuals_full_dt$Sequence, exclude = NULL)
```

**Other**

Number of index persons (with named partners) who identify as other:

```{r}
individuals_dt_of_other_who_named_partners <-
  individuals_dt[StudyID %in% IPs_reporting_partners_in_unified_dt & DemoRace == "Other",,]
nrow(individuals_dt_of_other_who_named_partners)
```


***How many partners are reported by the `r nrow(individuals_dt_of_other_who_named_partners)` other persons who named partners?***

```{r}
net_dt_of_other_who_named_partners <- net_dt[StudyIDFrom %in% individuals_dt_of_other_who_named_partners$StudyID,,] 

pt_counts_other_who_named_partners <- net_dt_of_other_who_named_partners[,.(N=.N),by=StudyIDFrom]

nrow(pt_counts_other_who_named_partners)
pt_counts_other_who_named_partners[,sum(N)]
pt_counts_other_who_named_partners[,summary(N)]
```

***Ans: `r pt_counts_other_who_named_partners[,sum(N)]`***.

***How many of these partners are in the individuals dataset, diagnosed, and sequenced?***

```{r}
idx <- which(net_dt_of_other_who_named_partners$StudyIDTo %in% individuals_dt$StudyID)
idx
other_reported_partners_in_individuals_dt <- net_dt_of_other_who_named_partners$StudyIDTo[idx] 
other_reported_partners_in_individuals_dt
unique(other_reported_partners_in_individuals_dt)
length(unique(other_reported_partners_in_individuals_dt))

other_partners_in_individuals_full_dt <- individuals_dt[StudyID %in% other_reported_partners_in_individuals_dt,,]
table(other_partners_in_individuals_full_dt$HIVDxDate, exclude = NULL)
table(other_partners_in_individuals_full_dt$HIVDxDate, exclude = NULL)
table(other_partners_in_individuals_full_dt$Sequence, exclude = NULL)
```

**Ethnicity = Hispanic**

Number of index persons (with named partners) who identify as Hispanic:

```{r}
individuals_dt_of_nonhispanic_who_named_partners <-
  individuals_dt[StudyID %in% IPs_reporting_partners_in_unified_dt & DemoHispanic == "True",,]
nrow(individuals_dt_of_nonhispanic_who_named_partners)
```


***How many partners are reported by the `r nrow(individuals_dt_of_other_who_named_partners)` other persons who named partners?***

```{r}
net_dt_of_nonhispanic_who_named_partners <- net_dt[StudyIDFrom %in% individuals_dt_of_nonhispanic_who_named_partners$StudyID,,] 

pt_counts_nonhispanic_who_named_partners <- net_dt_of_nonhispanic_who_named_partners[,.(N=.N),by=StudyIDFrom]

nrow(pt_counts_nonhispanic_who_named_partners)
pt_counts_nonhispanic_who_named_partners[,sum(N)]
pt_counts_nonhispanic_who_named_partners[,summary(N)]
```

**Ethnicity = Non-Hispanic**

Number of index persons (with named partners) who identify as Non-Hispanic:

```{r}
individuals_dt_of_nonnonhispanic_who_named_partners <-
  individuals_dt[StudyID %in% IPs_reporting_partners_in_unified_dt & DemoHispanic == "False",,]
nrow(individuals_dt_of_nonnonhispanic_who_named_partners)
```


***How many partners are reported by the `r nrow(individuals_dt_of_other_who_named_partners)` other persons who named partners?***

```{r}
net_dt_of_nonhispanic_who_named_partners <- net_dt[StudyIDFrom %in% individuals_dt_of_nonhispanic_who_named_partners$StudyID,,] 

pt_counts_nonhispanic_who_named_partners <- net_dt_of_nonhispanic_who_named_partners[,.(N=.N),by=StudyIDFrom]

nrow(pt_counts_nonhispanic_who_named_partners)
pt_counts_nonhispanic_who_named_partners[,sum(N)]
pt_counts_nonhispanic_who_named_partners[,summary(N)]
```

***Ans: `r pt_counts_nonhispanic_who_named_partners[,sum(N)]`***.

***How many of these partners are in the individuals dataset, diagnosed, and sequenced?***

```{r}
idx <- which(net_dt_of_nonhispanic_who_named_partners$StudyIDTo %in% individuals_dt$StudyID)
idx
nonhispanic_reported_partners_in_individuals_dt <- net_dt_of_nonhispanic_who_named_partners$StudyIDTo[idx] 
nonhispanic_reported_partners_in_individuals_dt
unique(nonhispanic_reported_partners_in_individuals_dt)
length(unique(nonhispanic_reported_partners_in_individuals_dt))

nonhispanic_partners_in_individuals_full_dt <- individuals_dt[StudyID %in% nonhispanic_reported_partners_in_individuals_dt,,]
table(nonhispanic_partners_in_individuals_full_dt$HIVDxDate, exclude = NULL)
table(nonhispanic_partners_in_individuals_full_dt$HIVDxDate, exclude = NULL)
table(nonhispanic_partners_in_individuals_full_dt$Sequence, exclude = NULL)
```