---
title: "R Notebook"
output: html_notebook
---

# Setup

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE,
                      root.dir = '~/code/rtp-network/out/')
library(data.table)
library(dplyr)
library(stringi)
library(network)
```

# Load the data

```{r}
eda_env <- readRDS("/users/akhann16/code/rtp-network/out/eda_objects.rds")

individuals_dt <- eda_env$individuals_dt
net_dt <- eda_env$net_dt 
named_pt_net <- eda_env$named_pt_net
named_pt_idx_005 <- eda_env$named_pt_idx_005 
named_pt_idx_015 <- eda_env$named_pt_idx_015
```

# Create edgelists from molecular clusters

Start by converting the unified individuals dataset
to an edgelist depending on cluster identity.  


```{r}
# Get a list of unique cluster IDs from the "ClusteredPhyloAny" column of the individuals_dt data frame
unique_clusters <- unique(na.omit(individuals_dt$ClusteredPhyloAny))

# Loop through the unique clusters and create an edgelist for each
edgelists <- list()
for (i in 1:length(unique_clusters)) {
  # Get the cluster ID
  cluster_id <- unique_clusters[i]
  
   # Print the cluster ID
  #print(paste("Processing cluster", cluster_id))
  
  # Filter the data frame to include only the rows corresponding to the current cluster ID
  df_cluster <- individuals_dt[individuals_dt$ClusteredPhyloAny == cluster_id, ]
  
  # Get a list of unique nodes in the current cluster
  nodes <- unique(df_cluster$StudyID)
  
  # Use combn() to get all combinations of nodes within the current cluster
  edges <- combn(nodes, 2) %>%
    t() %>%
    as.data.frame() %>%
    filter(V1 != V2) %>%
    distinct()
  
  # Rename columns
  colnames(edges) <- c("from", "to")
  
  # Store the edgelist in the list of edgelists
  edgelists[[i]] <- edges
}

# Combine all edgelists into a single edgelist
all_edges <- do.call(rbind, edgelists)

```

```{r}
# Get a list of unique cluster IDs from the "ClusteredPhyloAny" column of the individuals_dt data frame
unique_clusters <- unique(na.omit(individuals_dt$ClusteredPhyloAny))

# Loop through the unique clusters and create an edgelist for each
edgelists <- list()
for (i in 1:length(unique_clusters)) {
  # Get the cluster ID
  cluster_id <- unique_clusters[i]
  
  # Print the cluster ID
  print(paste("Processing cluster", cluster_id))
  
  # Filter the data frame to include only the rows corresponding to the current cluster ID
  df_cluster <- individuals_dt[individuals_dt$ClusteredPhyloAny == cluster_id, ]
  
  # Get a list of unique nodes in the current cluster
  nodes <- unique(df_cluster$StudyID)
  
  # Use combn() to get all combinations of nodes within the current cluster
  edges <- combn(nodes, 2) %>%
    t() %>%
    as.data.frame() %>%
    filter(V1 != V2) %>%
    distinct() %>%
    rename(source = V1, target = V2)
  
  # Add a column indicating the cluster ID for each edge, but only if the edges data frame is not empty
  if (!nrow(edges) == 0) {
    edges$cluster_id <- cluster_id
    # Store the edgelist in the list of edgelists
    edgelists[[i]] <- edges
  }
}

# Combine all edgelists into a single edgelist
all_edges <- do.call(rbind, edgelists)

```

```{r}
# Get a list of unique cluster IDs from the "ClusteredHIVTrace015" column of the individuals_dt data frame
unique_clusters <- unique(na.omit(individuals_dt$ClusteredHIVTrace015))

# Loop through the unique clusters and create an edgelist for each
edgelists <- list()
for (i in 1:length(unique_clusters)) {
  # Get the cluster ID
  cluster_id <- unique_clusters[i]
  
  # Skip to next iteration if cluster_id is empty
  if (cluster_id == "") {
    next
  }
  
  # Print the cluster ID
  print(paste("Processing cluster", cluster_id))
  
  # Filter the data frame to include only the rows corresponding to the current cluster ID
  df_cluster <- individuals_dt[individuals_dt$ClusteredHIVTrace015 == cluster_id, ]
  
  # Get a list of unique nodes in the current cluster
  nodes <- unique(df_cluster$StudyID)
  
  # Skip to next iteration if nodes is empty
  if (length(nodes) <= 1) {
    next
  }
  
  # Use combn() to get all combinations of nodes within the current cluster
  edges <- combn(nodes, 2) %>%
    t() %>%
    as.data.frame() %>%
    filter(V1 != V2) %>%
    distinct() %>%
    rename(source = V1, target = V2)
  
  # Add a column indicating the cluster ID for each edge
  edges$cluster_id <- cluster_id
  
  # Store the edgelist in the list of edgelists
  edgelists[[i]] <- edges
}

# Combine all edgelists into a single edgelist
all_edges <- do.call(rbind, edgelists)

```


Test:
```{r}
x <- as.numeric(table(individuals_dt$ClusteredHIVTrace015)[-1])
y <- lapply(x, function (x) choose(x, 2))
sum(unlist(y))

(sum(unlist(y)) == nrow(all_edges))
```

For now, let's use the HIV Trace 015 dataset. (Next round, I will add 
`phyloany` and `015`).

# Prepare data for regression analysis

How many index persons in the HIVTrace015 dataset?

```{r}
index_person_idx_in_015 <- which(individuals_dt$ClusteredHIVTrace015 != "")
index_persons_in_015 <- individuals_dt$StudyID[index_person_idx_in_015]
length(index_person_idx_in_015)

individuals_dt_w_index_persons_in_015 <- individuals_dt[index_person_idx_in_015,]
dim(individuals_dt_w_index_persons_in_015)
```

How many of these index persons are in the named partner list?

```{r}
named_pt_net%v% "vertex.names"
sequence015_persons_in_named_pt_net_idx <- which((index_persons_in_015) %in% (named_pt_net %v% "vertex.names"))

individuals_dt_w_index_persons_in_015$named_partner_indicator <- 0
individuals_dt_w_index_persons_in_015$named_partner_indicator[sequence015_persons_in_named_pt_net_idx] <- 1

individuals_dt_w_index_persons_in_015$named_partner_indicator <-
  as.factor(individuals_dt_w_index_persons_in_015$named_partner_indicator)

table(individuals_dt_w_index_persons_in_015$named_partner_indicator, exclude = NULL)
```

Run the regression model:

```{r}

table(individuals_dt_w_index_persons_in_015$DemoGender, exclude = NULL)
xtabs(~ factor(named_partner_indicator, exclude = NULL)+
      factor(DemoGender, exclude = NULL),
      data = individuals_dt_w_index_persons_in_015)
m1 <- glm(named_partner_indicator ~ DemoGender,
  data = individuals_dt_w_index_persons_in_015,
  family = "binomial")
summary(m1)

table(individuals_dt_w_index_persons_in_015$DemoRace)
xtabs(~ factor(named_partner_indicator, exclude = NULL)+
      factor(DemoRace, exclude = NULL),
      data = individuals_dt_w_index_persons_in_015)
m2 <- glm(named_partner_indicator ~ DemoRace,
  data = individuals_dt_w_index_persons_in_015,
  family = "binomial")
summary(m2)

table(individuals_dt_w_index_persons_in_015$SequenceSubtype)
xtabs(~ factor(named_partner_indicator, exclude = NULL)+
      factor(SequenceSubtype, exclude = NULL),
      data = individuals_dt_w_index_persons_in_015)
rowSums(xtabs(~ factor(named_partner_indicator, exclude = NULL)+
      factor(SequenceSubtype, exclude = NULL),
      data = individuals_dt_w_index_persons_in_015))
m3 <- glm(named_partner_indicator ~ SequenceSubtype,
  data = individuals_dt_w_index_persons_in_015,
  family = "binomial")
summary(m3)

table(individuals_dt_w_index_persons_in_015$SequenceSubtype)
m4 <- glm(named_partner_indicator ~ RiskMSM,
  data = individuals_dt_w_index_persons_in_015,
  family = "binomial")
summary(m4)

m5 <- glm(named_partner_indicator ~ RiskHRH,
  data = individuals_dt_w_index_persons_in_015,
  family = "binomial")
summary(m5)

m6 <- glm(named_partner_indicator ~ RiskIDU,
  data = individuals_dt_w_index_persons_in_015,
  family = "binomial")
summary(m6)

m7 <- glm(named_partner_indicator ~ HIVDxAge,
  data = individuals_dt_w_index_persons_in_015,
  family = "binomial")
summary(m7)

model <- glm(named_partner_indicator ~ DemoGender+DemoRace+SequenceSubtype,
  data = individuals_dt_w_index_persons_in_015,
  family = "binomial")
summary(model)
```

