---
title: "XYZ"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
```

## Introduction

Last time, we discussed the contact tracing dataset and my analysis thus far and questions on it. Vlad is interested in broadly knowing the factors of interest driving cluster growth: High VL, STI incidence, growing number of partners, etc.  To that end, we are interested in understanding if the 
changing number of partners over time can be measured, for instance, 
if we knew
the number of reported partners per person per year (named or unnamed) 
reported by the index cases (i.e., newly diagnosed persons). 

There are two other related questions:
  - If the recency of infection is [known]( https://docs.google.com/document/d/1j0uxP8P0DGefUEg8sr8ncT1GWIvz-6AOK42S8zcIans/edit#bookmark=id.yqp4tue4v2f).

  - A minor question about who we mean as ["Third Party"](https://docs.google.com/document/d/1j0uxP8P0DGefUEg8sr8ncT1GWIvz-6AOK42S8zcIans/edit#bookmark=id.un9xszbe7gwi) as the Notifier.
  
```{r}
data_dir <- "/gpfs/data/rkantor/rtp/datasets/D30_20211013_V1"
list.files(path=data_dir)
net_dt <- fread(paste0(data_dir, "/ContactTracingNetwork.csv"))
individuals_dt <- fread(paste0(data_dir, "/Individuals.csv"))

dim(net_dt)
str(net_dt)
sort(colnames(net_dt))
View(net_dt)

dim(individuals_dt)
colnames(individuals_dt)
```
  
  